---
title: "HW5: Markdown file"
author: "Kevin Liu"
date: "2023-11-12"
output: github_document
---

```{r include = FALSE}
library(tidyverse)
library(rvest)
```

# Problem 1

```{r }
raw_homicide_data =
  read_csv("./data/homicide-data.csv") |> 
  janitor::clean_names()
```

__Describe the raw data__

* There are `r nrow(raw_homicide_data)` rows and `r ncol(raw_homicide_data)` variables in the raw dataset.
* The 11 variables are "uid", "reported_date", "victim_last", "victim_first", "victim_race", "victim_age", "victim_sex", "city", "state", "lat", and "lon".
  * Respectively their variable types: chr, num, chr, chr, chr, chr, chr, chr, chr, num, num, and chr. 
  * The variables include full name of victim, their race, age, sex, the city and state and the coordinates of the homicide report and the date.

__Create a `city_state` variable__

```{r}
raw_homicide_data = 
  raw_homicide_data |> 
  
  #Create variable "city_state" which is a concatenation of city + state separated by , .
  mutate(city_state = paste(city, state, sep = ", "))
```

__Summarize__ 
...within cities to obtain the total number of homicides and the number of unsolved homicides.

```{r}
sum_homicides =
  raw_homicide_data |> 
  
  #Group by city_state 
  group_by(city_state) |> 
  
  #Summarize total_homicides as # of data inputs in each city and num_unsolved the # of data with "Closed without arrest" and "Open/No arrest" as the string value for disposition
  summarize(total_homicides = n(), num_unsolved = sum(disposition %in% c("Closed without arrest", "Open/No arrest"))) 
```

__For City of Baltimore, MD...__
...use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

```{r}
#Extract Baltimore_MD data from the summarized data aobve
baltimore_prop = 
  sum_homicides |> 
  filter(city_state == "Baltimore, MD")

#Extract the number of unsolved and total homicides from baltimore dataset
balt_unsolved = baltimore_prop$num_unsolved
balt_total = baltimore_prop$total_homicides

#Use prop.test function on the unsolved and total homicides from baltimore
balt_proptest = 
  prop.test(balt_unsolved, balt_total) |> 
  
  #Tidy the propotional test output
  broom::tidy()

#Extract the proportion and confidence interval from balt_proptes
balt_est = balt_proptest$estimate
balt_ci_low = balt_proptest$conf.low
balt_ci_high = balt_proptest$conf.high
```

__Now Run `prop.test` on each city...__
...in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

```{r}
#Create function that takes in city name and outputs proportions + CIs

#Iterate through all cities in the summarized data and produce df with proportions + CIs for each city

## Create list of city_states and their total_homicides + 
```

